# A Short Tour of the Predictive Modeling Process

```{r chapter-02-startup, include = FALSE}
library(tidymodels)
library(AppliedPredictiveModeling)
library(patchwork)

pkg_list <- c("patchwork", "AppliedPredictiveModeling", "tidymodels")
```

The R package used in this chapter are `r pkg_text(pkg_list)`.

The data are in several data frames contained in `r pkg(AppliedPredictiveModeling)`. The analysis in this chapter only uses the 2010 and 2011 data. 

```{r chapter-02-data}
data(FuelEconomy)
ls(pattern = "cars")
```

`cars2012` is not used in these analyses. 


## Working with the data

The relationship between engine displacement and fuel efficiency of all 2010 model year vehicles and new 2011 car lines: 

```{r chapter-02-fig-01}
cars2010 %>% 
 mutate(Year = "2010 Model Year") %>% 
 bind_rows(
  cars2011 %>% 
   mutate(Year = "2011 Model Year")
 ) %>% 
 ggplot(aes(x = EngDispl, y = FE)) + 
 geom_point(alpha = .3) + 
 facet_wrap(~ Year) + 
 xlab("Engine Displacement") + 
 ylab("Fuel Efficiency (MPG)")
```

## Resampling the training set


Create resampling folds for the 2010 data: 

```{r chapter-02-folds}
set.seed(121)
folds_2010 <- vfold_cv(cars2010)
```

## A simple linear regression model

Fit a linear regression model with a single linear term for engine displacement:

```{r chapter-02-linear-reg}
# Setup the model type and use `lm()` to fit the model: 
lm_spec <- linear_reg() %>% set_engine("lm")

# Create a control object that will save the out-of-sample predictions. 
ctrl <- control_resamples(save_pred = TRUE)

# Perform 10-fold cross-validation on the linear regression
lm_lin_resamples <- 
 lm_spec %>% 
 fit_resamples(FE ~ EngDispl, resamples = folds_2010, control = ctrl)

# Fit the linear model on the 2010 data
lm_lin_fit <- lm_spec %>% fit(FE ~ EngDispl, data = cars2010)

tidy(lm_lin_fit)
```

A function for creating plots of the predicted line and the observed versus predicted results: 

```{r chapter-02-diagnostic-function}
fuel_plots <- function(resamples, fit) {
 pred_line <- 
  predict(fit, cars2010) %>% 
  bind_cols(cars2010) %>% 
  arrange(EngDispl)
 
 profile_plot <- 
  cars2010 %>% 
  ggplot(aes(x = EngDispl, y = FE)) + 
  geom_point(alpha = .3) + 
  geom_line(data = pred_line, aes(y = .pred), col = "red") +
  xlab("Engine Displacement") + 
  ylab("Fuel Efficiency (MPG)")
 
 obs_vs_pred <- 
  collect_predictions(resamples) %>% 
  ggplot(aes(x = FE, y = .pred)) + 
  geom_abline(lty = 2) + 
  geom_point(alpha = .3) + 
  coord_obs_pred() + 
  xlab("Observed") + 
  ylab("Predicted")
 
 profile_plot + obs_vs_pred
}
```

 Quality of fit diagnostics for the linear regression model. The training set data and its associated predictions are used to understand how well the model works: 
 
```{r chapter-02-fig-02}
fuel_plots(lm_lin_resamples, lm_lin_fit)
```

The resampled RMSE: 

```{r chapter-02-linear-reg-rmse}
collect_metrics(lm_lin_resamples)
```

## A linear regression model with a quadratic term

Now add a quadratic term to the linear model. 

```{r chapter-02-linear-reg-quad}
lm_quad_resamples <- 
 lm_spec %>% 
 fit_resamples(FE ~ EngDispl + I(EngDispl^2), resamples = folds_2010, control = ctrl)

lm_quad_fit <- lm_spec %>% fit(FE ~ EngDispl + I(EngDispl^2), data = cars2010)
tidy(lm_quad_fit)
```

Quality of fit diagnostics for the quadratic regression model (using the training set):

```{r chapter-02-fig-03}
fuel_plots(lm_quad_resamples, lm_quad_fit)
```

Resampled RMSE goes down: 

```{r chapter-02-linear-reg-quad-rmse}
collect_metrics(lm_quad_resamples)
```

## Multivariate adaptive regression spline model

Tune a MARS model to the data, evaluating a model with two to five features: 

```{r chapter-02-mars, message = FALSE}
mars_spec <- 
 mars(num_terms = tune(), prune_method = "none") %>% 
 set_engine("earth") %>% 
 set_mode("regression")

ctrl <- control_grid(save_pred = TRUE)

mars_tune <-
 mars_spec %>%
 tune_grid(
  FE ~ EngDispl,
  resamples = folds_2010,
  control = ctrl,
  grid = tibble(num_terms = 2:5)
 )
```

The cross-validation profile for the MARS tuning parameter:

```{r chapter-02-fig-04}
autoplot(mars_tune, metric = "rmse")
```

Selecting the best tuning parameters: 

```{r chapter-02-mars-paramters}
# Show the top results:
show_best(mars_tune, metric = "rmse")

# Pick the numerically best:
mars_param <- select_best(mars_tune, metric = "rmse")
mars_param

# Update the model to substitute the value of `tune()` with
# the chosen value, then fit the model. 
mars_fit <- 
 mars_spec %>% 
 finalize_model(mars_param) %>% 
 fit(FE ~ EngDispl, data = cars2010)

# The fitted MARS terms:
cat(format(mars_fit$fit))
```

```{r chapter-02-mars-cuts, include = FALSE}
mars_cuts <- unique(mars_fit$fit$cuts[, "EngDispl"])
mars_cuts <- mars_cuts[mars_cuts > 0]
mars_cut_text <- knitr::combine_words(sort(mars_cuts))
```


Quality of fit diagnostics for the MARS model (using the training set). The MARS model creates several linear regression fits with change points at `r mars_cut_text` L:


```{r chapter-02-fig-05}
fuel_plots(mars_tune, mars_fit)
```

## Test set results

The test set data and with model fits for two models: 

```{r chapter-02-fig-06}
# Predict the test set using the two best models

lm_quad_test_res <- 
 lm_quad_fit %>% 
 predict(cars2011) %>% 
 bind_cols(cars2011) %>% 
 mutate(model = 'Quadratic Regression')

mars_test_res <- 
 mars_fit %>% 
 predict(cars2011) %>% 
 bind_cols(cars2011) %>% 
 mutate(model = 'MARS') 

lm_quad_test_res  %>% 
 bind_rows(mars_test_res) %>% 
 ggplot(aes(x = EngDispl, y = FE)) + 
 geom_point(alpha = .5) + 
 geom_line(aes(y = .pred), col = "red") +
 facet_wrap(~ model) +
 xlab("Engine Displacement") + 
 ylab("Fuel Efficiency (MPG)")
```
The test set statistics:

```{r chapter-02-test-stats}
lm_quad_test_res %>% rmse(FE, .pred)
mars_test_res %>% rmse(FE, .pred)
```
